%!TEX root = ../dissertation.tex

\begin{otherlanguage}{english}
\begin{abstract}
% Set the page style to show the page number
\thispagestyle{plain}
\abstractEnglishPageNumber

%(250 words)

%(objectives)
Generative Design (GD) allows architects to explore new kinds of design by using a programming-based approach.
Current GD environments are based on existing Computer-Aided Design (CAD) applications, such as AutoCAD or Rhinoceros 3D, which, due to their complexity, are slow and fail to give architects the feedback they need to explore GD.
In addition, current GD environments are limited by the fact that they need to be installed and, therefore, are not easily accessible in any computer.
Architects would benefit from a GD Integrated Development Environment (IDE) in the web that is accessible on any computer without installation, and, on the other hand, is at least as interactive as existing GD IDEs.

%(contents)
This thesis proposes a GD IDE based on web technologies.
Its main component is a web page, containing a program editing interface that allows the architect to make programs and view results in 3D.
To make the editing experience more intuitive, it also re-runs programs whenever they are changed, allows numeric literals to be adjusted by clicking and dragging, and highlights the relationship between program and results.
The IDE also includes a secondary application for exporting results to CAD applications installed in the architects computer.

%(results + conclusions)
With this approach, we were able to implement a GD environment that is accessible from any computer, offers an interactive editing environment, and integrates easily with the architect's workflow.
In addition, with respect to program running times, it has good performance that can be an order of magnitude faster than current GD IDEs.


% Keywords
\begin{flushleft}

\keywords{Generative Design, Web technologies, Integrated Development Environments, Architecture}

\end{flushleft}

\end{abstract}
\end{otherlanguage}
